Interoception Scale - Data Cleaning

### 2024-11-29 study1/analysis/1_cleaning.qmd
Update 1_cleaning.qmd
@AnafNeves @EmmaLBenn @maisiebennett added another outlier detection check + streamlined the code a bit + added some TODOs

```{r}
#| warning: false
#| code-fold: true
#| error: false

@@

# df <- read_csv("C:/Users/maisi/OneDrive/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

# df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

df <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study1/data/rawdata_participants.csv")

## Attention Checks

```{r}
#| warning: false
#| code-fold: true
#| error: false

df <- df |>
  dplyr::mutate(
    A1 = dplyr::case_when(
      # "I always know that I am attentively doing a study"
      Sexual_State_A == 6 | Sexual_State_A == 5 ~ "Pass",
      TRUE ~ "Fail"),
    A2 = dplyr::case_when(
      # "Even if I am anxious, I should now answer all the way to the left"
      Anxious_SkinThermo_A == 0 ~ "Pass",
      TRUE ~ "Fail"),
    A3 = dplyr::case_when(
      # "I often experience sensations, and I will answer zero to this question"
      Nociception_ColonBladder_A == 0 ~ "Pass",
      TRUE ~ "Fail"),
    A4 = dplyr::case_when(
      # "In general, I am very sensitive and attentive to the questions I am       currently answering"
      Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5 ~ "Pass",
      TRUE ~ "Fail"),
    A5 = dplyr::case_when(
      # "I often pay attention to the answers I am giving"
      Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5 ~ "Pass",
      TRUE ~ "Fail"),
    A6 = dplyr::case_when(
      # "I can always accurately answer to the left on this question to show that I am reading it" 
      Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1 ~ "Pass",
      TRUE ~ "Fail"),
    A7 = dplyr::case_when(
      # "I can always accurately perceive that to this question I should answer the lowest option"
      Accuracy_Genital_A == 0 ~ "Pass",
      TRUE ~ "Fail"),
    A8 = dplyr::case_when(
      #  "Sometimes I notice that I need to answer all the way to the right"
      Confusion_ColonBladder_A == 6 ~ "Pass",
      TRUE ~ "Fail"),
  )

# Create colum with the total number of Failed attention checks 
df$Total_Attention_Failed <- apply(df[, c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "A8")], 1, function(x) sum((x == "Fail")))

p_fail <- df |>
  ggplot(aes(x = Total_Attention_Failed)) +
  geom_bar(aes(fill = factor(Total_Attention_Failed))) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5)),
    axis.title.x = element_blank()
  )

p_fail

```{r}

removed_participants <- nrow(df |> filter(Total_Attention_Failed >= 1))

df <- df |>
  filter(Total_Attention_Failed == 0)
```

`r removed_participants` participants were removed for having failed at least 1 attention check.

## Experiment Enjoyment

```{r}
#| warning: false
#| code-fold: true
#| error: false

@@

@@

```{r p_demographics_age_gender}
#| warning: false
#| code-fold: true
#| error: false

@@

```{r p_demographics_education}
#| warning: false
#| code-fold: true
#| error: false

@@

```{r p_demographics_ethnicity}
#| warning: false
#| code-fold: true
#| error: false

@@

```{r p_demographics_country}
#| warning: false
#| code-fold: true
#| error: false

@@

```{r p_demographics_map}
#| warning: false
#| code-fold: true
#| error: false

@@

```{r p_summary}
#| warning: false
#| code-fold: true
#| error: false

@@

```{r}
#| warning: false
#| code-fold: true
#| error: false

>>>>

```{r}
#| warning: false
#| code-fold: false
#| error: false

@@

df <- read.csv("../data/rawdata_participants.csv")
```

## Feedback

### Experiment Enjoyment

```{r}
#| warning: false

#| error: false

@@

**TODO: Add feedback comments table**

## Exclusions

```{r}
outliers <- list()
```

### Attention Checks

```{r}
#| warning: false
#| error: false

dfchecks <- df |>
  dplyr::mutate(
    # "I always know that I am attentively doing a study"
    A1 = ifelse(Sexual_State_A == 6 | Sexual_State_A == 5, 0, 1),
    # "Even if I am anxious, I should now answer all the way to the left"
    A2 = ifelse(Anxious_SkinThermo_A == 0, 0, 1),
    # "I often experience sensations, and I will answer zero to this question"
    A3 = ifelse(Nociception_ColonBladder_A == 0, 0, 1),
    # "In general, I am very sensitive and attentive to the questions I am currently answering"
    A4 = ifelse(Sensitivity_Cardiac_A == 6 | Sensitivity_Cardiac_A == 5, 0, 1),
    # "I often pay attention to the answers I am giving"
    A5 = ifelse(Sensitivity_Gastric_A == 6 | Sensitivity_Gastric_A == 5, 0, 1),
    # "I can always accurately answer to the left on this question to show that I am reading it"
    A6 = ifelse(Accuracy_Respiratory_A == 0 | Accuracy_Respiratory_A == 1, 0, 1),
    # "I can always accurately perceive that to this question I should answer the lowest option"
    A7 = ifelse(Accuracy_Genital_A == 0, 0, 1),
    # "Sometimes I notice that I need to answer all the way to the right"
    A8 = ifelse(Confusion_ColonBladder_A == 6, 0, 1),
    .keep = "none"
  ) 
dfchecks$Total <- rowSums(dfchecks)

dfchecks |>
  mutate(Total = as.factor(paste0(Total, "/8"))) |> 
  ggplot(aes(x = Total)) +
  geom_bar(aes(fill = Total)) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Failed Attention Checks", y = "Number of Participants", subtitle = "Number of failed attention checks per participant") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.title.x = element_blank(),
  )
```

```{r}
outliers$attentionchecks <- df$Participant[dfchecks$Total >= 1]
```

We removed `r length(outliers$attentionchecks)` (`insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for having failed at least 1 attention check (out of 8).

### Mahalanobis Distance

```{r}
# Compute distance
dfoutlier <- performance::check_outliers(select(df, contains("_Q")), 
                                         method=c("lof"),
                                         threshold=list(lof = 0.0001)) |> 
  as.data.frame() |> 
  mutate(Participant = fct_reorder(df$Participant, Distance_LOF),
         Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
         Outlier = ifelse(Outlier_LOF == 1, "Distance Outlier", "Valid"),
         Outlier = ifelse(Outlier_AttentionCheck == 1, "Failed Attention Check", Outlier))

# parameters::parameters(glm(Outlier_AttentionCheck ~ Distance_LOF, data = dfoutlier, family = "binomial"))

outliers$distance <- as.character(dfoutlier[dfoutlier$Outlier_LOF == 1, "Participant"])
outliers$distance <- outliers$distance[!outliers$distance %in% outliers$attentionchecks]

dfoutlier |> 
  ggplot(aes(x=Participant, y=Distance_LOF)) +
  geom_bar(aes(fill=Outlier), stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

**TODO: Add a little paragraph explaining what it is etc.**
We removed `r length(outliers$distance)` (`insight::format_percent(length(outliers$distance) / nrow(df))`) participants based on multivariate distance.

```{r}
df <- filter(df, !Participant %in% outliers$attentionchecks)
df <- filter(df, !Participant %in% outliers$distance)
```

@@

```{r p_demographics_age_gender}
#| warning: false

#| error: false

@@

```{r p_demographics_education}
#| warning: false

#| error: false

@@

```{r p_demographics_ethnicity}
#| warning: false

#| error: false

@@

```{r p_demographics_country}
#| warning: false

#| error: false

@@

```{r p_demographics_map}
#| warning: false

#| error: false

@@

```{r p_summary}
#| warning: false

#| error: false

@@

```{r}
#| warning: false

#| error: false
